Appendix C

### 2025-07-04 supplemental/appendix_C.qmd
manuscript update with revisions after review

---
title: 'Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections'
# subtitle: 'The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections'
shorttitle: 'Effect of Recruiting Military Veterans on Confidence in 2024 US Elections' # aog-article-pdf
shortauthors: Isaiah      # aog-article-pdf
abstract: |
  I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment support the notion that emphasizing veterans as the target of election worker recruitment efforts eases pre-election insecurity. Expectations of electoral fraud and concerns for voter safety were lower among those who read an announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate. In addition, concerns over voter safety and potential for violence were lower among those who received the treatment.
authors:
  - name: Isaiah Espinoza
    email: gespinoz@umd.edu
    roles: author
    affiliation:
      - name: University of Maryland
        department: Department of Government and Politics
        city: College Park
        state: MD
        country: US
reference-section-title: References
knitr: 
  opts_chunk: 
    dev: "ragg_png" # same as knitr::opts_chunk$set(dev = "ragg_png")
editor_options: 
  chunk_output_type: console
---

@@

# load data
load("data/research-paper-data-20241029.Rdata")

@@

```{r}
#| label: fun-likert-plot

# custom function used to create Likert Plots

# To avoid excessive copy/paste, I made a function to create Likert plots
# it is mostly a wrapper around `ggstats::gglikert`, but with a specific set up
# particular colors, theme, etc. 
likert_plot <- function(data, x, ...,  
                        symmetric = FALSE,
                        variable_labels = NULL,
                        label_size = 2.5,
                        vline = FALSE,
                        title = waiver(), 
                        subtitle = waiver(),
                        caption = waiver(),
                        xlab = waiver()){

  p <- ggstats::gglikert(
    data = data, include = {{ x }}, ...,
    variable_labels = variable_labels,
    symmetric = symmetric)+

    # customize color
    ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
    # custom theme
    theme_bw(base_family = "TeX Gyre Pagella")+
    theme(
      legend.position = 'bottom',  # place legend on bottom
      axis.title.y.left = element_blank(), # blank out y-axis label
      axis.text.x = element_blank(), # remove percentage text along x-axis
      strip.text.y.right = element_text(angle = 0) # make facet label horizontal
        )

    if (vline==TRUE) {
    p <- p + geom_vline(xintercept = 0, color = 'black', linewidth = 1.2)
  } else {
      p
  }

  p <- p + ggplot2::labs(..., title = title, subtitle=subtitle, caption=caption, x = xlab)
  p

}

```{r}
#| label: fun-barplot

# function for creating barplots
custom_barplot <- function(data, x, group, title = NULL, subtitle = NULL,
                           caption = NULL, legend_fill = NULL,
                           xlab = NULL){
  b <- data |> 
    group_by({{ group }}, {{ x }}) |> 
    count() |> 
    drop_na() |> 
    group_by({{ group }}) |> 
  mutate(prop = round(n/sum(n), digits = 3),
         pct = prop*100,
         res = str_c(pct,'% (', n, ')', sep = "")) |> 
  ggplot(aes(x = {{ x }}, y = pct, fill = {{ group }}))+
  geom_bar(position = 'dodge', stat = 'identity')+
  geom_text(aes(label = res), position = position_dodge(1.0), size = 2.5, vjust = -0.7)+
  scale_fill_grey(start = 0.5, end = 0.1)

  b <- b + ggplot2::labs(
    title = title,
    subtitle = subtitle,
    caption = stringr::str_wrap(caption, width = 99),
    fill = legend_fill,
    y = "Percentage",
    x = stringr::str_wrap(xlab, width = 85))
  b
}

```{r}
#| label: fun-logit-gof

# custom function to return a certain selection of logistic regression model
# goodness-of-fit statistics
lrm.gof.stats <- function(model, ...){

  out <- data.frame(
    nobs = stats::nobs(model),
    null.deviance = model$null.deviance,
    df.null = model$df.null,
    logLik = as.numeric(stats::logLik(model)),
    deviance = stats::deviance(model),
    df.residual = stats::df.residual(model),
    chisq = as.numeric(model$null.deviance - stats::deviance(model)),
    df = as.numeric(model$df.null - stats::df.residual(model)),
    'P(>chi)' = pchisq(q=model$null.deviance - stats::deviance(model),
                       df = model$df.null - stats::df.residual(model), 
                       lower.tail = F)
  )
  out <- dplyr::as_tibble(out) |>  
  # rename column variables
  dplyr::rename(
    "Num.Obs."      = "nobs",
    "Log.Lik"       = "logLik",
    "Deviance"      = "deviance",
    "Deviance Null" = "null.deviance",
    "DF"            = "df",
    "chisq"          = "chisq",
    "P(>chisq)"     = "P..chi."
  )
  return(out)
}

{{< pagebreak >}}

## Introduction

Election administration officials make efforts to sustain public trust and confidence in the fairness and accuracy of elections, and attempt to boost such confidence where it may be deprived. Concerns for safety have developed among election staff and voters in more recent elections. Regular measures are taken to enhance the *trustworthiness* of the electoral process through practices meant to improve the conduct, transparency, or overall administration of elections in the United States.

Although election officials undertake great efforts to enhance the *trustworthiness* of election administration, public *trust*[^1] in elections is a psychological construct influenced by many things outside of election official control such as partisanship or elite rhetoric [@hooghe2018; @sances2015]. Moreover, a person's evaluation of the election in hindsight is often influenced by the election outcome itself [@daniller2019; @stewart2022]. Thus, measures taken by election officials can be undermined, trivialized, or made irrelevant depending on how one feels after the election results have come out.

Such volatile attitudes and evaluations post-election can leave a lasting impression that election officials must contend with upon the next election cycle [@levendusky2024; @bowler2024]. For instance, we have witnessed many people's outright refusal to accept the 2020 U.S. election results as legitimate despite consistent review of the evidence confirming the results as fair and accurate. Such a case demonstrates that public trust in elections is, at best, only partial to trustworthiness of election administration in the United States.

One point of contention that election officials have faced in the past regard evaluation of election workers. Previous literature has focused on how voter interaction with election workers [@claassen2008], voting technology [@herrnson2009], and the voter experience generally [@atkeson2007], influences evaluations of election administration. As such, election worker competency has been examined as a factor significant to evaluations of performance of elections [@hall2007; @hall2009]. However, considering that individual perceptions and preconceived notions play a huge role in cognition [@cikara2014; @vanbavel2021], it is reasonable to expect that the group an election worker hails from would be an important influence upon the voter's evaluation of the electoral process.

Supposing such is the case, we can expect that information about *who* (i.e., which groups) election officials are targeting in publicized recruitment efforts would lessen particular election insecurity, and concurrently, boost confidence. That is to say, it is reasonable to expect that telling people *who* will be working and volunteering as election staff would ease election insecurity, and therefore improve confidence that the election will be conducted fairly, accurately, and safe for all involved.

I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment show that emphasizing veterans as the target of election worker recruitment efforts is associated with greater confidence in elections and lower pre-election insecurity. Expectations of electoral fraud and concerns for voter safety were lower among those who read an announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned[^2]. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate. In addition, concerns over voter safety and potential for violence were lower among those who received the treatment.

This paper is structured as follows. First, I provide a brief background on public trust in election administration. I synthesize a review of relevant literature with a focus on how political and social science has conceptualized and ascertained public trust in elections. Next, I supply reasoning for why military veterans are singled out as the relevant subset of the population expected to influence confidence in elections. I provide testable hypotheses of the study before moving on to describe the design of the survey, the measurement instruments therein, and the conceptual and operational definitions of the variables of interest. I explain my reasoning and method for constructing the primary dependent variable, which I broadly refer to as *confidence in elections*. I close by offering my interpretation of the results and suggest potential avenues for future inquiry.

## Background: Election Administration and Public Confidence

Election officials have tried hard to inspire confidence in the administration and conduct of elections by improving the degree to which elections are trustworthy. Development and implementation of procedures such as post-election auditing of ballots and logic-and-accuracy testing of ballot tabulation equipment are prominent examples adding to the long history of efforts to enhance the trustworthiness of election administration in the United States.

Prior to the year 2000, one of the main issues facing election administration was recruiting enough election workers to volunteer at the polls (i.e., poll workers) [@maidenberg1996]. Election worker recruitment is still much of an issue in the current era as it was then, perhaps worse [@ferrer2024]. In addition to ensuring election offices were adequately staffed, the controversy of the 2000 general election made the public more attentive to issues concerning the conduct and administration of elections. In particular, voting technology [@herrnson2009] and election worker competence was of interest in election studies [@claassen2008; @hall2009; @hall2007]. Following the passage of the Help America Vote Act in 2002, election officials efforts to boost public confidence in the conduct and administration of elections revolved primarily around the accuracy of vote counts, ballot tabulation equipment or voting machines, the commitment of election staff, and more [@atkeson2007].

In 2024, election officials made valiant efforts to boost public confidence elections within an intensified political climate that appeared quite hostile to election officials [@edlin2024; @brennancenterforjustice2024]. Although polling around the time indicated that most people thought that U.S. elections would be run at least somewhat well [@nadeem2024], many election officials nationwide took efforts to assuage the worry of those most skeptical.

Election anxiety was high in the lead up to the 2024 elections in the United States. Concerns for voter safety and the prospect of political violence remained prescient and compelled many local officials to prepare for the worst [@edlin2024; @doubek2024]. Election officials in Washoe County, Nevada, installed panic buttons for election staff that would alert a monitoring center to summon law enforcement [@lincoln2024]. Nevada also passed a law making it a felony to harass, threaten, or intimidate election workers [@nevadasecretaryofstate2023]. Leading up to election day, news outlets reported that election work had become a seemingly dangerous job [@wire2024]. A Brennan Center survey report stated that, "...large numbers of election officials report having experienced threats, abuse, or harassment for doing their jobs" [@edlin2024]. Concerns over the fairness of elections and accuracy of vote counts intensified, heightening concerns over the prospect of political violence and, in turn, increased worry for the safety of voters and election workers alike. Since election worker performance is significant to public evaluation of elections, added safety concerns that drive out election staff and repel volunteers can only detract from trustworthiness of the institution.

## Literature Reivew

Assessing the public's *trust* in elections has not been straightforward. Inquiry into public trust in elections has been approached by scholars of political science in many different ways [@cook2005], often distinguishable by the scope of the research question and more or less constrained by the particular conception of public trust. Quite often, trust in election administration is conflated with trust in government writ large, government legitimacy, government or system responsiveness, or even satisfaction with democracy [@daniller2019]. At this level, not only is the level of public trust in elections sometimes vague, but there's little consideration over the difference between such attitudes pre-election and post-election. In contrast, a considerable amount of research tends to conceive of public trust in accordance with the institution in question [@atkeson2007; @hooghe2018].

Election administration^[It is important to point out here that election administration is just one institution that interacts with many other institutions and processes which form the electoral system *writ large* [@stewart2022]. The way in which I discuss the institution of election administration, as well as the way in which I discuss and distinguish public trust in elections from broader conceptions of public trust, is borrowed substantively from @stewart2022's conceptual framework of the same. Distinctions between *trust* and *trustworthiness* and the notion of election administration as an institution are already apparent inspirations convenient for situating the theory of this research inquiry.] is just one part of the larger set of institutions which form the electoral system. As such, the performance of the institution along with the rest "...lends credibility to the outcome of an election: whether it is considered by citizens and the international community to be fair and legitimate." [@stewart2022, 236].

Trust in the conduct of elections concerns aspects of elections that fall squarely within the institution of election administration. At this level, for instance, public trust in elections is ascertained by capturing assessments about the perceived accuracy of vote counts (e.g., whether votes are/were counted as intended).

Intuitively, enhancing public trust in elections would best be accomplished by enhancing the *trustworthiness* of the institution, e.g., enhancing vote tabulation equipment and technology, conducting audits. However, trust and confidence in elections has become ever more precarious over the last few election cycles. Especially considering public polling data since 2000 shows that confidence that votes were, or would be, counted as intended was in a consistent decline despite efforts towards bolstering election integrity and trustworthiness [@sances2015]. This is even more pronounced considering the role that partisanship has had on such confidence over accuracy of vote counts [@sances2015; @stewart2022].

There's also stark difference in public trust before the election has occurred compared to after, a phenomenon referred to as the "winner-loser gap"; the "winners" are those who supported the winning candidate and the "losers" are those who supported the losing candidate. Much research has been dedicated to analyzing the sentiment of electoral winners vs losers, and vice versa [@nadeau1993; @daniller2019]. Opinions of electoral trust gathered after the election has occurred, however, are limited considering the well-recognized impact that the electoral outcome itself has on feelings of public trust in elections [@daniller2019].

As such, it is questionable whether we can characterize public trust in the pre-election period as the same trust after the election results have come out. The former is *anticipatory*---i.e., the kind that is more or less anxious given the uncertainties surrounding the election. The latter is *empirical*---a judgement discerned in hindsight after the experience of the election event has occurred. As discussed by @stewart2022, not only was confidence in vote count accuracy influenced by the voter experience, but so too were evaluations of election officials [-@stewart2022, 242-243]. Not to mention the influence that the election results would also have on such evaluations. In other words, evaluations of trust considered in hindsight are influenced by the voter experience and the outcome of the election results. This study, in contrast, focuses primarily on that *anticipatory* kind of confidence, which speaks more to those insecurities based on perceptions of the institution's trustworthiness than upon the particular voter experience.

Regardless of the measures taken by election officials to boost public confidence in the *trustworthiness* of election administration (e.g., conducting audits, testing election machines), public *trust* and confidence in elections more generally is apt to shift dramatically post-election based on factors such as partisanship, elite rhetoric, particular state policies, and more [@nadeau1993; @carter2024; @coll2024a]. Moreover, prior research has found that evaluation of election workers themselves are an important factor when it comes to levels of public confidence in the electoral process [@hall2007; @hall2009; @claassen2008]. Such studies focused on the quality of the voter experience with reference to the interaction between voter and election worker.

Beyond the general competence of election workers, however, the quality of the voter experience may be influenced merely by prior impressions about *who* comprises election staff and volunteers. Political and other social science researchers have recognized for some time the power that group identity can have over attitudes and perception [@xiao2016; @xiao2012; @vanbavel2021]. As such, we can expect that information identifying the particular groups being recruited to serve as election staff will be enough to improve confidence in elections, and lessen expectations of electoral fraud. The next section elaborates on the proposed theory, terms utilized, and why military veterans are of particular interest in this regard.

# Theory

This theory is constrained to public confidence in the administration of elections in the United States. To be clear, the scope conditions in which this theory applies concern public trust and distrust in election administration in the anticipatory period prior to the election event. Moreover, this study is, to a limited extent, indirectly related to public regard about veteran members of the U.S. Armed Forces.

## Trust and Distrust in Elections

The theoretical relationship between trust and distrust is important to understand, especially when it comes to the dependent outcome of interest. Before going further, it is necessary to explicate the conception *confidence in elections* with respect to trust and distrust.

Trust is, after all, merely *belief* held in the face of uncertainty. Trust is reliance on the veracity of belief in the face of uncertainty. Said another way, trust is assured belief dependent on its supposed veracity; a state of certainty with respect to some belief or judgement. It is "assured" because it is one's belief that relies on something rather than nothing. Hence quantitative methods of the sciences are replete with things such as *confidence* intervals around our estimates. Therefore, I understand trust to refer to the state or quality of certainty around one's beliefs or expectations; the extent to which one is assured of their expectations regarding what is or what will be the case.

The kind of *trust in elections* I refer to concerns the expectations drawn up from one's judgement about the functioning of the electoral process---the extent to which one is assured in the belief that the electoral process will be fairly and accurately administered. I follow the notion of public trust in elections explicated by @stewart2022,

> "Public trust in U.S. election institutions comes down to whether voting machines accurately record votes, voter registration systems accurately record those eligible to vote, geographic information systems accurately assign voters to voting districts, election-night reporting systems accurately aggregate and communicate election results to officials and the public, and postelection audit and canvassing procedures proceed impartially and in accordance with the law." [-@stewart2022, 236]

That being said, I also account for expectations of electoral fraud, which I refer to as *distrust* in elections distinguished from trust. Distrust in this sense is also a belief, an expectation, about the anticipated functioning of the process. Note, however, that the prefix *dis-* in distrust implies "apart", "lack of, not" or "opposite". Conventionally, distrust is defined simply as the absence of trust, thus denoting a simple inverse and mutually exclusive relationship. However, as it concerns the psychology of an individual, *a lack of trust doesn't necessarily imply opposite expectations or opposing beliefs*. In particular, a lack of trust that votes will be accurately recorded, etc., doesn't necessarily imply inversely proportional expectations that electoral fraud will occur. In other words, although the relationship between trust and distrust is inverse, this relationship is not mutually exclusive.

A person who is certain that electoral fraud will occur clearly lacks trust in elections; an expectation that electoral fraud will occur necessarily implies a lack of trust in elections regardless of positive expressions of trust. However, the reverse isn't also true; expressing a lack of trust in elections *doesn't necessarily imply* an expectation that electoral fraud will occur. Rather, a lack of trust implies an absence of assurance in the belief that the process will be conducted fairly and accurately---a lacking quality of certainty around said belief. An absence of such assurance denotes a state of *insecurity*. Other sources of uncertainty may challenge a person's confidence regarding the integrity of elections independent from anything that would lead them to believe electoral fraud will occur or has occurred. However my use of distrust refers to the assured belief that the outcome will be invalid---that outcomes will be independent of the process as prescribed, thus rendering outcomes as arbitrary. Specifically, distrust implies assurance in the belief that electoral fraud will occur.

As such, trust and distrust represent opposing degrees of confidence along the same spectrum. However, despite the inherent contrast, one can hold both positive degrees of trust and distrust, feeling relatively confident in both respects. Accordingly, *equivalent degrees of trust and distrust cancel out*, rendering one relatively more insecure about their expectations of the future despite the fact that they may report feeling confident in either direction respectively. When trust and distrust are considered reflective of one's confidence about future expectations, then both can be placed along the same spectrum. Therefore, a lack of confidence denotes insecurity.

This means that a person's baseline level of confidence in elections must take into account their present degree of distrust---i.e., their expectations of electoral fraud. Something may cause a person to have greater expectations that election fraud will occur, which in turn will always lower their trust in elections. In contrast, a person's trust in the integrity of the electoral process is reduced by the extent to which they expect that electoral fraud will occur.

Thus, it follows that a person distrusts the integrity of the electoral process to the extent that they are to some extent certain that electoral fraud will occur or has occurred. They are, perhaps, poised to discount election results as invalid or illegitimate in advance. Either way, both trust and distrust refer to some positive degree of confidence (i.e., assurance) regarding one's expectations. They are distinguished by reference to either the normative idea about what *should* come about, or what is *supposed* to come about given the logical structure and integrity of the process. If the structure and integrity of some process is coherent and secure, respectively, then the outcome that follows will be valid. Should the integrity of that process be damaged or opaque, then the process becomes vulnerable and validity of the results are open to challenge. With respect to the administration of U.S. elections, the confident voter expects that votes will be counted as voters intend, election staff will competently administer elections, election technology will be secure from nefarious tampering, and that the process will be fair for all involved.

## Why Veterans

Election officials don't focus on recruiting people of a particular demographic, status, class, or prestige in order to shore up public trust in elections, and they don't discriminate recruitment efforts based on who the public thinks is best suited to secure elections from fraud.

Election officials are likely to be agnostic as to who dedicates their time to civil service such as election work. Especially since staffing has been an issue since at least the 1990s [@maidenberg1996; @ferrer2024]. In 2020, such efforts were made far more difficult by the COVID-19 pandemic [@mena2020; @abbate2020a]. There's no special reason to target veterans for recruitment above other groups. Indeed, there's no reason to discriminate recruitment efforts at all if the point is purely to fill staffing vacancies.

Veteran military service members are, after all, civilians; Civilians with shared (but varied) experience and status of prior military service, but who are no more or less qualified for this civil service^[Research is limited and mixed on whether veterans in general are more inclined, however [@nesbit2011].] than non-veteran civilians. To be frank, announcements about concerted efforts to recruit veterans as election workers should not reasonably influence one's trust that votes will be counted more accurately, or that election fraud will not occur; there is no known evidence demonstrating that actually employing veterans as election staff or volunteers further improves the safety, security, and overall integrity of elections in the United States. Moreover, there's no known evidence of positive influence on public perceptions of said outcomes when voters encounter military veterans working as staff or volunteers at elections sites.

Rather, the claim that recruiting veterans to work as election staff and volunteers would improve public confidence in elections emerges as non-profit organizations---mainly comprised of veteran and military family members themselves---operate under, and promote, that assumption [@looker2024; @lawrence2024; @wentling2024]. Non-profit groups of late have made incredible strides in targeted recruitment efforts under the notion that veterans are uniquely capable, committed, and best fit to serve the needs of the United States as civil servants and volunteers at election sites. Indeed, this study can be seen as a test of a particular assumption that veterans have about themselves---the assumption that their mere presence and civic engagement will inspire confidence in elections among the public.

There was a sudden push to target veterans for recruitment efforts that arose shortly after the events on Capitol Hill on January 6th, 2021. After the 2020 election, large efforts were made to recruit military veterans and their families to work or volunteer as election staff [@wetheveterans2022; @nflfootballoperations2022]. Prior to that point, young people were sometimes given special mention as targets of election worker recruitment efforts [@powerthepolls2020; @herndon2020]. Generally, however, recruitment efforts cast a wide net, indiscriminate of who applies [@ross2020; @conde2020].

One can speculate that the motivation to associate military veterans with civic engagement and democracy may be intended to counter negative perceptions and impressions given by the proportion of veteran service members arrested for taking part in the events on January 6th [@milton2021; @loewenson2023; @jensen2022a]. Especially with regard to research demonstrating that willingness to support violent efforts to overturn election results (in support of Trump) is, on average, more common among veterans than among matched samples of non-veterans [@pape2024]. This is in addition to a strengthened association portrayed in media outlets between military veterans and militias [@steinhauer2020]. Prior research has substantiated such a connection between veterans and militia groups. @cooter2024 notes from her 3-year ethnographic fieldwork among Michigan militia members that, "...approximately 40% of militia leaders and 30% of members had previous military experience. Most of these veterans actively sought out such groups, as opposed to being recruited by them" [see also @cooter2013]. Thus, countering such associations by promoting a different image of veterans to the mass public and veterans alike seems like a reasonable motivation. Yet such speculation is just that.

The key findings of literature on confidence in elections hardly point to public regard for veterans as a mechanism for the theoretical expectations of improving overall confidence in elections and improving feelings of voter safety at election sites. Comparable research finds that confidence in elections is increased through elite rhetoric from fellow partisans affirming legitimacy of elections [@clayton2023], or through the provision of information about the electoral process or other official messaging of election officials themselves [@gaudette2025]. Neither of the studies just cited, however, consider attitudes or sentiment about veterans as a source for dispelling distrust and improving trust in elections. To be clear, there is no immediate "puzzle" in the literature that immediately draws out public sentiment for military veterans as a source for bolstering confidence in elections. That being said, the general public perception, attitudes, or even stereotypes about military veterans are significant to consider.

It is important to make clear two considerations that underpin the notion that information about veterans will improve trust and quell distrust in elections. First, positive public regard for veterans *in general* is sufficient to elicit greater trust and lower distrust in elections despite the fact that information about veterans as election workers, or even merely information on efforts to recruit veterans, has no bearing on the actual integrity of the electoral process. An underlying assumption here is that there *is* something special about veterans, but this doesn't include a proposition on exactly *what* that special something might be in theory. This is partly due to the limited research available on public regard for veterans, specifically in terms of elucidating an underlying mechanism for such positive favor. Moreover, there is limited research on public opinion about former and active military service members distinguished separately from public opinion about the military as an institution in general. Endorsement or support of government policies intended to benefit veterans specifically (e.g., federal funding on veteran healthcare) are easy to conflate as interchangeable with public attitudes about veterans as a group in particular. This, in turn, permits general expectations of positive influence due to the public's overarching positive regard for veterans, but disallows for expectations regarding how respondents would actually process and interpret a treatment that only mentions efforts to recruit them for election work. In short, we are sure that people sustain positive views for veterans generally, but beyond that I make no special effort to propose and examine why that is the case in greater detail.

The evidence that can be gathered concerning public regard for military veterans shows that, generally, veterans are perceived favorably among the general public; although military recruitment shows a downward trend as of late, public perceptions of veterans are overwhelmingly positive [@kleykamp2023]. In my review of the literature, I found only one study which assessed public opinion about veterans specifically and examined whether attitudes differed by group [@kleykamp2018]. @kleykamp2018 investigated the extent to which positive regard for military veterans in terms of both social distance and deservingness varies across different groups, and the extent to which such support is exaggerated. Their research found that levels of support for veterans tended to be predictably lowest, but also most exaggerated, among those who were younger and non-White. Determining what groups hold special favor for veterans still leaves it unclear as to what it is about veterans that engenders greater confidence in elections. Yet, generally, all groups expressed exaggerated support for veterans in more than one way [@kleykamp2018], suggesting that public regard for veterans is prescribed by widespread social norm.

This is taken together with literature suggesting that public trust in elections is malleable to factors, experiences, events, or phenomena that range in logical relevance to the actual structure or administration of elections. On the one hand, public trust is, indeed, influenced by information specifically about the processes of conducting elections [@gaudette2025]. On the other hand, the public's trust in elections is not beholden to the demonstrated *trustworthiness* of the institution of election administration. Thus the first consideration is simply that the positive public regard for veterans in general may function sufficiently as a cue to elicit greater trust in elections despite relevance.

Second, veterans are a particularly potent group where mention of one's veteran status seems to have a positive, calming, or nullifying effect on attitudes. More so than just the general positive favor afforded to veterans, the group appears to have a somewhat pacifying effect on negative or otherwise contemptible discrimination, prejudice, and social stigma. For instance, recent research shows that, during his campaign in the 2020 Democratic Primaries, Pete Buttigieg's military background mitigated discrimination against him when he was presented as a veteran married to a man [@magni2024]. Similar research has also found that a candidate's veteran status affords them better evaluations regarding competency in particular issue areas (e.g., war competence) [@teigen2013; @hardy2019]. Moreover, veteran status seems to mitigate or nullify usual stigmas associated with mental illness. That is to say, there is negative stigma associated with mental illness [@corrigan2002] and such stigma incurs labor market discrimination [@hipes2016], but evidence suggests that veteran status overrides such stigma and discrimination [@maclean2014]. Or, in another light, mental illness is seemingly more *understandable* (i.e., permissible) for veterans given the presumptive reasons for their mental strife. And media framing as such plays an important role on public perception [@kleykamp2015]. Relatedly, attitudes about homelessness and PTSD---ranging from perceived causes of homelessness, whether the federal government should dedicate more resources, perceived effectiveness of policies, to compassion for homeless individuals---are generally more favorable for the homeless population among veterans than for the homeless population of general adults [@tsai2021]. Such effects primarily pertain to attitudes about individual people or groups (e.g., a political candidate, prospective employees, the homeless population). The capacity of one's veteran status to pacify prejudice, discrimination, and stigma likely extends out to quell negative and disfavorable attitudes regarding institutions to a similar extent, e.g., distrust.

Given the heightened tensions and electoral anxieties, the expectations surrounding the unique pacifying effect that veterans have on attitudes about individuals or other groups is extended to determine whether a similar effect is found when the target of one's consideration is not just a person or group but the institution of election administration itself. Thus the second consideration for "why veterans" are expected to dispel distrust in elections pertains to pacifying effect of negative or disfavorable attitudes observed in relevant literature.

The stimulating influence of veterans as a group in the abstract is expected to counteract distrust and insecurity pertaining to the institution of elections administration, as well as inspire greater confidence in the same. There's scant research or evidence available to postulate a well-formed theory on the mechanism underlying the expected effects of veterans as a group on attitudes about institutions and processes, rather than on attitudes about persons or groups. What may be posited at best, however, is that distrust and insecurity in elections among the public is expectedly reduced to some extent through generally positive public regard for military veterans and their unique capacity to quell disfavorable attitudes.

## Hypotheses

The outcome of interest is *confidence in elections*, which refers to trust in elections---i.e., expectations that elections will be administered fairly and accurately---compensated by distrust in elections---i.e., expectations that electoral fraud will occur. Considering the generally positive perception afforded to veterans by the public, it is reasonable to suspect that efforts that promote recruitment of veteran military service members to work as election staff and volunteers would boost public *confidence in elections administration*. Indeed, that is the primary hypothesis of this study. Although citizens may vote in various ways across the country (e.g., by mail, ballot drop box, in-person), the simple announcement that election officials are engaging in efforts to recruit veterans to work as election staff may boost confidence in elections administration regardless of how, or whether, an individual plans to vote. Formally,

> H~1~: Announcements that election officials are recruiting miilitary service veterans to work as election staff and volunteers will be associated with greater confidence in elections administration compared to announcements that do not mention military veterans.

Asking participants whether they supported programs intent on recruiting anyone to work as election staff would garner support regardless of the target group. Considering that support and admiration for veterans is generally high among the population [@kleykamp2023], then discerning the impact on one's confidence would require a survey experiment designed to determine whether publicized recruitment efforts targeting veterans as a group would have any special effect on confidence in elections administration.

In addition, prior research has shown that confidence that one's own vote would be counted as intended has been much stronger than confidence that votes nationwide would be treated likewise [@sances2015; @stewart2022]. In other words, the public is more trusting of elections closer to home. Taking into account that confidence is lower for elections conducted outside of one's local area, then information that veterans are being recruited to help conduct elections in locations beyond one's locale may close this gap between confidence in elections within one's local area and confidence in elections elsewhere. Although it may be difficult to discern whether recruiting veteran service members would improve confidence in elections near one's local area in light of such hometown favoritism, a positive effect on confidence in elections held beyond one's local area---where confidence is already expectedly lower---can help discern the impact of the group alone on public opinion. Therefore, any observed disparity of confidence in elections between those within one's local area and those beyond will be smaller among those who are presented with information about election official recruitment efforts of veteran service members. Formally,

> H~2~: Differences in confidence in elections in one's local area and confidence in elections in Maricopa County, AZ, will be smaller among those who are presented with an announcement that election officials are recruiting miilitary service veterans to work as election staff and volunteers compared to those presented with an announcement of recruitment efforts that do not mention military veterans.

Since the results and events following the 2020 election loomed large in anticipation of the U.S. 2024 general elections, it is reasonable to take into account beliefs about legitimacy of the 2020 election. This group who denies the legitimacy of the 2020 election results were possibly the most apprehensive (distrusting) of the 2024 election. However, given that veteran service members are held in high regard, generally speaking, announcing that veterans are being actively recruited to work as election staff and volunteers is expected to positively influence confidence in elections among those who remained firm in the belief that the 2020 election results were illegitimate. That is to say, despite their belief about the illegitimacy of the 2020 election results, I expect that informing them about veteran recruitment efforts to improve their confidence in elections. I take into account the relationship between partisanship and 2020 election legitimacy beliefs and include partisanship as a control in the analysis[^3].

> H~3~: Among those who refute the legitimacy of the 2020 election results, those presented with an announcement that election officials are recruiting veteran service members for election jobs will be associated with greater levels of confidence in elections compared to those presented with an announcement of recruitment efforts that do not mention military veterans.

Prior research has found that political candidates benefit from having a military background [@mcdermott2015], especially evaluations of candidate's prospective performance on particular war issues (e.g., foreign policy, terrorism) [@hardy2019]. The idea is that prior military service provides meaningful information to use as a heuristic in regard to such "war issues". Although veterans are no longer actively serving, there's potentially some lingering general assumptions about the capacities and motives of military veterans to provide safety and security arising from the nature of their prior occupation as members of the U.S. armed forces. Insofar as the physical safety of people at election sites is a "military matter", then the potential that election sites may be staffed with military veterans may ease related concerns about the safety and security of those places. As such, I expect that concerns about potential for violence will be lower, and confidence in voter safety higher, when presented with information that veterans are being actively recruited to work and volunteer in election offices.

> H~4a~: Concerns about the potential for violence, threats of violence, or intimidation while voting in person will be lower among those presented with an announcement that election officials are recruiting veteran service members for election jobs compared to those presented with an announcement of recruitment efforts that do not mention military veterans.

> H~4b~: Confidence that voters will be safe to vote in-person at election sites will be higher among those presented with an announcement that election officials are recruiting veteran service members for election jobs compared to those presented with an announcement of recruitment efforts that do not mention military veterans.

## Experiment Design and Survey Measures

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve confidence in elections and ease insecurity, a recent experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment ($n = 650$) control ($n = 637$) conditions[^4].

The median age was $46$ (mean age was $47$), $51.13\%$ ($n = 658$) women, $46.46\%$ ($n = 598$) men, and approximately $1.24\%$ who identified as either Non-binary/third gender ($n = 7$) or preferred not to say ($n = 9$). The sample primarily identified as White or Caucasian $75.76\%$ ($n = 975$), while all other non-White respondents comprised $23.08\%$ ($n = 297$) of the sample. Those who held a graduate level degree (e.g., Master's, Doctorate, or Professional level) comprised $12.9\%$ ($n = 166$) of the sample; those with either degree at the Associate or Bachelor's level comprised $35.82\%$ ($n = 461$), while $21.99\%$ ($n = 283$) had some college but no degree; and $28.13\%$ ($n = 362$) had either a high school level or equivalent education or less than high school. The largest proportion of the sample were Democrat at $43.98\%$ ($n = 566$), followed by Republicans at $41.96\%$ ($n = 540$). The proportion of true Independents^[Independent 'leaners' were grouped into the respective party affiliation in which they lean.] was $12.59\%$ ($n = 162$). Among the sample, $34.11\%$ ($n = 439$) said they had an immediately family member who was currently serving or had previously served in the U.S. military; $64.72\%$ ($n = 833$) didn't have a family member who served. Approximately $8.24\%$ ($n = 106$) of the sample were veterans, while $1.17\%$ ($n = 15$) reported to be actively serving.

>>>>

---
title: "Appendix C"
subtitle: "Polychoric Item and Score Correlations of Trust and Distrust"
format:
  typst: 
    toc: false
    keep-typ: true
    papersize: us-letter
    mainfont: TeX Gyre Pagella
    fontsize: 11pt
execute: 
  echo: false
knitr: 
  opts_chunk: 
    dev: "ragg_png" # same as knitr::opts_chunk$set(dev = "ragg_png")
---

@@

# load data
load(here::here("data", "research-paper-data-20241029.Rdata"))

# set ggplot theme that will be used for every plot
# theme_get()
old <- theme_set(theme_bw()) # capture current theme

# set/update new theme
ggplot2::theme_set(
  new = ggplot2::theme_bw(base_family = 'TeX Gyre Pagella')+
    ggplot2::theme(legend.position = 'bottom'))

@@
